Glyph-maps for visualizing climate data and models

### 2011-09-05 paper.tex
working on scaling section

Graphical displays benefit from a \emph{visual reference grid} \citep{cleveland:1993a} to compare values mapped to position along a line. The same is true for glyph-maps, but it is not feasible to make a full set of axies for each icon. A complete axis would dominate the picture, and draw the eye away from the pattern of the data. The types of comparisons that are needed across icons are:

@@

\section{Partitioning patterns into shape and size}~\label{sec:scale}

@@

@@

>>>>

Graphical displays benefit from a \emph{visual reference grid} \citep{cleveland:1993a} to compare values mapped to position along a line. The same is true for glyph-maps, but it is not feasible to make a full set of axes for each icon. A complete axis would dominate the picture, and draw the eye away from the pattern of the data. The types of comparisons that are needed across icons are:

@@

Glyph-maps using trend lines enable the reading of trends across
spatial locations directly from the plot. In comparison, to read trend
from facetted colored maps one needs to spot the difference between
the different maps. A more direct comparison would be to make a single
map with the slope of the trend coded as color. Cleveland's
\citep{cleveland:1993a} perceptual tests suggests that the reader can
perceive the positional difference of the lines better than the color
differences. Accurate perception of color differences is difficult.

\section{Effect of scale}~\label{sec:scale}

@@

*** Global/local min/max, remove intercept/average, data vs model scale

Scaling of the data can produce dramatically different emphases on
structure. For spatiotemporal data the most natural scale is to use
the global minumum and maximum. The resulting icons will have size
relative to the magnitude of the full range of the data. Locations
with overall large or small temperature will be the focus. Figure
\ref{fig:glyph-map} shows data scaled to -1 to 1 based on the global
minimum and maximum, using:

\begin{equation}
y_{\amaj} = \frac{y_{\amaj}-y_{min}}{y_{max}-y_{min}} \times 2 - 1
\end{equation}

\noindent where $y_{min} = \mbox{minimum}\{y_{st}; s=1, \dots S, t=1, ..., T \}, y_{max} = \mbox{maximum}\{y_{st}; s=1, \dots S, t=1, ..., T\}$, $S=$ is the number of spatial locations and $T=$ is the number of time points. *** Reference other plots that use global scale too.

Alternatively, scaling locally at each spatial location enables
exploring the seasonal scale, or shape differences, from one location
to another. Fig \ref{***} uses local scaling using:

\begin{equation}
y_{\amaj} = \frac{y_{\amaj}-y_{s,min}}{y_{s,max}-y_{s,min}} \times 2 - 1
\end{equation}

\noindent where $y_{s, min} = \mbox{minimum}\{y_{st}; t=1, ..., T \}, y_{s,max} = \mbox{maximum}\{y_{st}; t=1, ..., T\}$.

@@

*** Connection to small multiples

*** Scale of icons relative to data or model range
